\chapter{Taste of category theory}
\section{Categories}
As you have had training in sets and functions, we are able to introduce some category theory that will quickly become useful - most of the objects in mathematics form a
category and this language will be extremely convenient to find similarities between different branches of mathematics.
As we remember, it is not possible to create a set of all sets without having a contradiction. So let's use a word \textbf{collection} of sets or \textbf{class}
\footnote{Formal treatment of classes - collections that are in some sense bigger than sets - is introduced in von Neumann–Bernays–Gödel and Morse-Kelley set theories.}
 of sets - that is not a set and we don't know how to express it formally - but what has an intuitive sense.

\begin{definition}
  A \textbf{category} is:
  \begin{enumerate}
    \item a collection of objects such that
    \item for each pair objects $A$, $B$ in the collection there is a \emph{set} $\Hom(A,B)$ called the set of \textbf{morphisms} or \textbf{maps} of \textbf{arrows} such that
    \item for morphisms $f\in \Hom(A,B), g\in \Hom(B,C)$ there is a morphism $g\circ f\in \Hom(A,C)$. We also require:
      \begin{itemize}
        \item that composition of morphisms is associative: $h\circ(g\circ f)=(h\circ g)\circ f$, where $h\in \Hom(C, D)$
        \item we have morphisms $\Id_X: X\to X$ for every object $X$ such that $f\circ\Id_A = f=\Id_B\circ f$ for $f\in \Hom(A,B)$
      \end{itemize}
  \end{enumerate}
  If $f\in \Hom(A,B)$, we can also write $f:A\to B$ or $A\xrightarrow{f} B$. Some authors also write $\text{Mor}(A,B)$ for $\Hom(A,B)$ and $gf$ for $g\circ f$. If the collection of objects happens
  to be a set, we call it \textbf{small category}.
\end{definition}

\begin{example}
  We already know very well a category - the category of sets and functions. Let's check carefully that is actually is a category:
  \begin{enumerate}
    \item objects are just sets
    \item take $\Hom(A,B)$ as a set of all functions from $A$ to $B$ (why is it a set?)
    \item define composition of morphisms just as function composition
      \begin{itemize}
        \item composition of functions is associative (recall why)
        \item identity morhpism is just identity function of a set
      \end{itemize}
  \end{enumerate}
\end{example}

\begin{exercise}
  Consider a category with one singleton: $\{\{0\}\}$, where $\{0\}$ is the only object, and functions as arrows. How many arrows are in this category?
\end{exercise}

\begin{exercise}
  Consider a category with two singletons: $\{\{0\}, \{1\}\}$, and functions as arrows. How many arrows can be in this category? Hint: 4 numbers.
\end{exercise}

\subsection{Morphisms}
In set theory we introduced special functions - injections, surjections and bijections. We would like to generalise it into category theory. Unfortunately, we cannot just state their properties in terms
of elements, as objects do not need to be sets and we cannot take a look at elements. Therefore, let's introduce the following:

\begin{definition}
  Let $f: A\to B$ be a morphism. We say that is it \textbf{left-cancellative} or that it is a \textbf{monomorphism} iff for every two morphisms $g,g' \in \Hom(X,A)$ we have:
  $$f\circ g = f\circ g'\Rightarrow g=g'.$$
  Sometimes monomorphism is written as $f: A\hookrightarrow B$. We will also refer to monomorphism as \textbf{monic} morphisms\footnote{Note that some authors distinguish between monomorphisms and monic
  morphisms.}.
\end{definition}

As you can prove, this works as injections in the category of sets:

\begin{exercise}
  Here you should prove that in the category of sets, "injective" are identical to "monic".
  \begin{enumerate}
    \item Prove that an injection in the category of sets is a monomorphism.
    \item Prove that if function $f: A\to B$ is not injective, then it is not monic. Hint: if $f(x)=f(y)$ for $x\neq y$, create functions from $\{x,y\}$ to $A$ showing that $f$ is not monic.
  \end{enumerate}
\end{exercise}

You can see that if we considered a different category, with less morphisms, the construction wouldn't work. So even if we consider sets and functions, but with some restrictions or additional structure,
we need to carefully investigate the relation between monomorphisms and injections.

\begin{exercise}
  Consider a set $\{\{0,1\}, \{10,11\}\}$. Construct three morhpisms such that this set becomes a category, but there is a non-injective monomorphism.
\end{exercise}

\begin{definition}
  Let $f: A\to B$ be a morphism. We say that is it \textbf{right-cancellative} or that it is a \textbf{epimorphism} iff for every two morphisms $g,g' \in \Hom(B,Y)$ we have:
  $$g\circ f = g'\circ f\Rightarrow g=g'.$$
  We will also refer to epimorphisms as \textbf{epic} morphisms\footnote{Note that some authors distinguish between epimorphisms and epic morphisms.}.
\end{definition}

\begin{exercise}
  Here you should prove that in the category of sets, "surjective" are identical to "epic".
  \begin{enumerate}
    \item Prove that a surjection in the category of sets is an epimorphism.
    \item Prove that if function $f: A\to B$ is not surjective, then it is not epic. Hint: if $f$ is not surjective, then both $\Image f$ and $B\setminus \Image f$ are non-empty. Define suitable functions.
  \end{enumerate}
\end{exercise}

Also here a problem may appear - there are categories in which surjections make sense, but are not identical to epimorphisms.

\begin{exercise}
  Create a category with a non-surjective epimorphism.
\end{exercise}

Also we have an object looking as bijection:

\begin{definition}
  Let $f: A\to B$ be a morphism. We say that is is an \textbf{isomorphism} if there is a morphism $g:B\to A$ such that $f\circ g=\Id_B$ and $g\circ f=\Id_A$.
\end{definition}

\begin{exercise}
  Prove that in any category an isomorphism is both monic and epic.
\end{exercise}

\begin{exercise}
  Create a category with a morphism that is both monic and epic, but is not an isomorphism.
\end{exercise}

But there are nice catgories, as the category of sets, for which it holds.

\begin{exercise}
  Prove that in the category of sets, isomorphisms are just bijections.
\end{exercise}

\section{Functors and natural transformations}
