\section{Cardinality}
\subsection{Finite sets}
\begin{definition}
  The \textbf{cardinality} $|X|$ of a finite set $X$ is defined as the number of elements in $X$.
\end{definition}

\begin{example}
  Let $A=\{0,1,2,3\}$. Then $|A|=4$.
\end{example}

\begin{exercise}
	What is the cardinality of $\{a, a+1, a+2, \dots, a+n\}$?
\end{exercise}

\begin{theorem}
  \textbf{Inclusion-exclusion principle}
  If $X$ and $Y$ are finite sets, then:
  $$|X\cup Y|=|X|+|Y|-|X\cap Y|.$$
\end{theorem}

Intuitively, adding two sets we count elements in each set twice and then subtract the number of elements that were counted twice. The formal proof goes as follows:

\begin{exercise}
  Prove the inclusion-exclusion principle:

  \begin{enumerate}
    \item Let $X$ and $Y$ be finite, disjoint (that is $X\cap Y=\emptyset$) sets. Prove that:
    $$|X\cup Y| = |X| + |Y|.$$
    \item Prove that for $A\subseteq X$, where $X$ is finite, we have $|X\setminus A|=|X|-|A|$. \hint{$X\setminus A$ and $A$ are disjoint and sum up to $X$...}
    \item Prove that $$|X\cup Y|=|X|+|Y|-|X\cap Y|$$ for finite sets $X,\, Y$ (now we don't assume that they are disjoint). \hint{What is $(X\setminus(X\cap Y))\cup Y$?}]
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Prove that if $B\subseteq A$, and $A$ is finite, then $|B|\le |A|$. When does the equality hold?
\end{exercise}

\begin{exercise}
  Prove that $|\mathcal P(A)|=2^{|A|}$ for a finite set $A$. Do you see why the power set $\mathcal P(A)$ is often referenced as $2^A$?
\end{exercise}

\begin{exercise}
    Let $A,B,C$ be finite sets. Prove that:
		$$|A\cup B\cup C| = |A|+|B|+|C| - |A\cap B| - |B\cap C|-|C\cap A| + |A\cap B\cap C|.$$
\end{exercise}

\begin{exercise}
  Let $X=\{1,2,\dots, 2018\}$.
\end{exercise}

\subsection{Characteristic functions}
\begin{definition}
  Fix a set $U$. For each subset $A\subseteq U$ we define it's \textbf{characteristic function} or \textbf{indicator function} as:

  $$1_A: U\to \{0,1\}$$
  $$1_A(x) = \begin{cases}1, \text{ if } x\in A\\ 0, \text{ if } x\notin A\end{cases}$$
\end{definition}

\begin{example}
  Consider a set $U$. Then $1_\emptyset(x)=0$ and $1_U(x)=1$ for every $x\in U$. It's usually abbreviated as:
  $$1_\emptyset=0, 1_U=1.$$
\end{example}

\begin{exercise}
  Let $A,B\subseteq U$. Prove that:
  \begin{enumerate}
    \item $1_{A_\cap B}=1_A\cdot 1_B$\footnote{It means that for every $x\in U$ we have $1_{A_\cap B}(x)=1_A(x)\cdot 1_B(x)$}
    \item $1_{A^c}=1-1_A$, where $A^c=U\setminus A$
    \item $1_{A\cup B}=1_A+1_B-1_A\cdot 1_B$
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Prove inclusion-exclusion principle for finite sets using characteristic functions. \hint{Write $1_{A\cup B}$ in terms of $1_A, 1_B, 1_{A\cap B}$ and sum its values over
  all elements in \textit{finite} set $A\cup B$.}
\end{exercise}

\subsection{Comparing cardinalities}
Although we feel comfortable in counting elements of \textit{finite} sets, we don't know how to say how to compare infinite sets - there is no natural number we could use to denote
their cardinalities!

Therefore, we'll try another approach. Assume that we have a set of children and a set of toys. If we want to compare them, we can either try to calculate how many children and toys there are (it may be very hard if there are lots of children and lots of toys) or to ask each child to get one toy. If every child has \textit{one} toy and no toys are left, we know
that there are exactly as many children as toys! We'll use this approach to compare infinite sets.

\begin{definition}
  Let $A$ and $B$ be two sets. If there exists a bijection $f:A\to B$, we say that $|A|=|B|$ (are of the same cardinality).
\end{definition}

\begin{example}
  $|\N|=|2\N|$, where $2\N$ is a set of all even natural numbers, as we can find a bijection $n\mapsto 2n$. It's a surprising result, as $2\N\subseteq \N$ is a \textit{proper} subset. If $\N$ was finite, all it's proper subsets would have smaller cardinalities!
\end{example}

\begin{exercise}
  Being of the same cardinality has similar properties to these of equivalence relation\footnote{... but as there is no sets of all sets, it \emph{is not} formally an equivalence relation.}. Prove that:
  \begin{enumerate}
    \item $|A|=|A|$
    \item $|A|=|B|$ implies that $|B|=|A|$ \hint{Bijections have inverses.}
    \item if $|A|=|B|$ and $|B|=|C|$, then $|A|=|C|$ \hint{Composition of bijections is a bijection as well.}
  \end{enumerate}
\end{exercise}

\begin{exercise}
  By considering the function
  $$(-1,\,1) \ni x \mapsto \frac{x}{1-x^2}\in \R,$$
  prove that \emph{every} non-empty open interval $(a,\, b)$ has the same cardinality as $\R$. \hint{Can you find a bijection between $(a,\, b)$ and $(-1,\,1)$?}
\end{exercise}

\begin{definition}
  We say that a set $X$ is \textbf{countably infinite} if $|X|=|\N|$. Usually we'll write that $\aleph_0:=|\N|$ (read "aleph 0").
  We say that a set $X$ is \textbf{countable} if $X$ is finite or countably infinite.
\end{definition}

\begin{example}
  Sets $\N, 2\N, \{0,1,6, 41\}$ are countable.
\end{example}

\begin{example}
  $\Z$ is countable: $0\mapsto 0, 1\mapsto 1, 2\mapsto -1, 3\mapsto 2, 4\mapsto -2,\dots$
\end{example}

\begin{exercise}
  Prove that a subset of a countable set is countable.
\end{exercise}

\begin{exercise}
  Let $A$ and $B$ be countable sets. Prove that $A\cup B$ and $A\cap B$ are countable.
\end{exercise}

\begin{exercise}
  Let $A$ and $B$ be countable. Prove that $A\times B$ is countable. \hint{You can write all elements of $A$ as $a_1,a_2,\dots$ and the elements of $B$ as $b_1, b_2,\dots$.
  Think about an ordering $(a_1,b_1); (a_1, b_2), (a_2, b_1); (a_1, b_3), (a_2, b_2), (a_3, b_1); \dots$ (some terms may be repeated if $A$ and $B$ are not disjoint, think how to fix it).}
\end{exercise}

\begin{exercise}
  Prove that $\Q$ is countable.
\end{exercise}

\begin{exercise}
  Let $\mathcal A$ be a countable family of countable sets. Prove that $\bigcup \mathcal A$ is countable.
\end{exercise}

\begin{exercise}
  Prove that is $X$ is an infinite set, then it contains a countably-infinite subset $S\subseteq X, |S|=\aleph_0$.
\end{exercise}

The last exercise shows that we can compare cardinalities. That is, if we can find a bijection between $A$ and \textit{some subset} of $B$, we can be sure that $B$ contains at least
as many elements as $A$. This is exactly requiring the existence of an \textit{injection} from $A\to B$.

\begin{definition}
  If there exists an injection $f:A\to B$ we say that $B$ has greater or equal cardinality than $A$ and write $|A|\le |B|$. If $|A|\le |B|$ and $|A|\neq |B|$, we write
  $|A| < |B|$ (that is: we can find an injection from $A$ to $B$, but there is no bijection between them).
\end{definition}

% \begin{exercise}
% 	Above we find the way of saying that two cardinalities are equal using existence of a bijection. Let's find a way to compare which is less using
% 	another kind of function.
% 	\begin{enumerate}
% 		\item Let $O_n=\{1,2,\dots,n\}.$ Prove that there is no injection from $O_{n+1}$ into $O_n$. Hint: use mathematical induction.
% 		\item Let $A$ and $B$ be finite. Prove that there is an injection from $A$ to $B$ iff $|A| \le |B|.$
% 	\end{enumerate}
% \end{exercise}
%
% \begin{exercise}
% 	Prove in one: if there is an injection from $A$ onto $B$ and an injection from $B$ into $A$, then there exists a bijection from $A$ onto $B$. Hint: you know that $|A|\le B$ and $|B|\le |A|$.
% \end{exercise}

As compositions of injections is an injection, we have the following:

\begin{exercise}
	Let $A,\,B$ and $C$ be sets. Prove that if $|A|\le |B|$ and $|B|\le |C|$, then $|A|\le |C|$.
\end{exercise}

In the following exercise, you will prove that there are more reals that natural numbers.

\begin{exercise}
	We define $X=\{x\in \R : 0\le x\le 1\}$.\footnote{We also need to choose one
	 convention of writing reals. Consider a number with finite expansion, e.g. 0.123. We can write this either as 0.123(0) or 0.122(9).}
	Prove that there is no surjection from $\N$ onto $X$ (injection is easy to find...). \hint{Assume that you have written all the elements of $X$ in a single column. Can you construct a real number that does not occur in the list? It should differ from the first number at least at one digit, and same for other numbers.}
\end{exercise}

\begin{exercise}
	We know that $|\R| > |\N|$. Using binary expansion prove that $|\R|=|\mathcal P(\N)|$. Do you see similarity between the previous result and $2^n > n$ for natural $n$?
\end{exercise}

We may suspect the following:

\begin{theorem}
\textbf{Cantor's theorem} $|A| < \left|\mathcal P(A)\right|$ for any set $A$.
\end{theorem}

\begin{exercise}
	Prove Cantor's theorem in two steps:
  \begin{enumerate}
    \item Find an injection from $A$ to $\mathcal P(A)$.
    \item To prove that there is no surjection, you can consider any $f:A\to \mathcal P(A)$ and a set
    $$X=\{a\in A : a\notin f(a)\}\in \mathcal P(A).$$
    Is there $x\in A$ for which $f(x)=X$?
  \end{enumerate}
\end{exercise}

\begin{exercise}
	Use Cantor's theorem to prove that there is no set of all sets.
\end{exercise}

The last exercise is a bit more complicated, although it's statement looks rather obvious:

\begin{theorem}
	\textbf{Cantor-Schroeder-Bernstein theorem} If $|A|\le|B|$ and $|B|\le |A|$, then $|A|=|B|$.
\end{theorem}

We will prove this repearing the proof of Banach and Tarski. The proof is tricky, so it was split into several parts.

\begin{definition}
  We say that a function $F: \mathcal P(A)\to \mathcal P(B)$ is \textbf{monotonous} if $F(X)\subseteq F(Y)$ whenever $X\subseteq Y$.
\end{definition}

\begin{theorem}
  \textbf{Knaster-Tarski fixed point theorem} A monotonous function $F: \mathcal P(A)\to \mathcal P(A)$ posseses a fixed point, that is: there exists a set $S$ such that $F(S)=S$.
\end{theorem}

\begin{exercise}
  Prove the fixed point theorem by considering a set:
  	$$S=\bigcup \mathcal U, \text{~where~} \mathcal U = \{Y\in \mathcal P(A) : Y\subseteq F(Y)\}.$$
    \hint{
      Each $Y$ is a subset of $S$, use the monotonicity to conclude that $Y\subseteq F(Y)\subseteq F(S)$ and $S=\bigcup \mathcal U\subseteq F(S)$.
      To prove the other inclusion, use $F(S)\subseteq F(F(S))$, what shows that $F(S)\in \mathcal U$.
    }
\end{exercise}

\begin{definition}
  Recall that if $f: A\to B$ and $X\subseteq A$, then we define
    $$f[X]=\{b\in B : b=f(x) \text{ for some } x\in X\}.$$
\end{definition}

\begin{exercise}
  \begin{enumerate}
    \item Let $f: A\to B$ and $g:B\to A$ be injections. Prove that the function
      $$F:\mathcal P(A)\to \mathcal P(A),~F(X)=A\setminus g[B\setminus f[X]]$$
      has the monotonicity property.

    \item Let $S$ be the fixed point of $F$. Prove that $A\setminus S\subseteq \text{Im}\,g$.
    \item Conclude that there exists $\tilde g: A\setminus S \to B$ such that $g(\tilde g(a))=a$ for $a\in A\setminus S$.
    \item Prove that function
      $$h(x) =
        \begin{cases}
          f(x), & x\in S\\
          \tilde g(x), & x \in A\setminus S
        \end{cases}
       $$
       is a bijection from $A$ to $B$, what ends the proof of the Cantor-Schroeder-Bernstein theorem.
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Recall that every non-empty open interval $(a,\,b)$ has the same cardinality as $\R$. Prove that every closed interval $[a,\,b]$ with $b > a$ has the same cardinality as $\R$. 
\end{exercise}
